---
title: "Report 1_gruzlica"
author: "Joanna Neubauer"
date: "2024-10-28"
output:
  html_document:
    toc: true
    toc_float: true
---

## Setup
During setup, libraries:

- devtools
- dplyr 
- ggplot2
- EDWAR

are installed and necessary data is downloaded and cached.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r setup_libs, echo=FALSE, message=FALSE}
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
library(dplyr)
library(ggplot2)
theme_set(theme_minimal())
devtools::install_github("rstudio/EDAWR")
library(EDAWR)
```

```{r setup_cache, echo=FALSE, message=FALSE}
df <- tb
```


## Data Summary

Data consists of 3 800 records specifying country,	year,	sex and number of sick people for every age group(child, adult,	elderly). 

The period of time the data covers is from 1995 to 2013. 

100 countries are taken into consideration.


```{r data_summary }
knitr::kable(summary(df))
```

#### Example Data
```{r data_summary_head }
knitr::kable(head(df, 10))

length(unique(df$country))
```


## Number of sick people per gender

Below summary shows the number of sick people, based on their sex and regardless of their age. It takes information though all the years and from all the countries.

```{r sick_per_gender, echo=FALSE, message=FALSE}
changed_df <- df %>%
  select(-(country:year)) %>%
  group_by(sex) %>%
  summarise(summed = sum(child, na.rm = TRUE) + sum(adult, na.rm = TRUE) + sum(elderly, na.rm = TRUE))

knitr::kable(changed_df)

ggplot(changed_df, aes(x = sex, y = summed, fill = sex)) +
  geom_col() 
```


## Sickness every year

Data is shown for children, adults and elderly people over the years.

```{r sickess_by_year, echo=FALSE}
sickness_by_year_df <- df %>%
  select(-country) %>%
  group_by(year) %>%
  summarise(sum_child = sum(child, na.rm = TRUE),
            sum_adult = sum(adult, na.rm = TRUE),
            sum_elderly = sum(elderly, na.rm = TRUE))

knitr::kable(sickness_by_year_df)

ggplot(sickness_by_year_df, aes(x = year)) +
  geom_line(aes(y = sum_child, color = 'Child')) +
  geom_line(aes(y = sum_adult, color = 'Adult')) +
  geom_line(aes(y = sum_elderly, color = 'Elderly')) +
  labs(y = "No of sick people")
```

## Sickness every year by country

The data is shown for single country, picturing number of sick people over the years. All with age group distinction.

```{r sickness_by_year_by_country, echo=FALSE, message=FALSE}
countries <- unique(df$country)

sickness_by_year_df <- df %>%
  group_by(country, year) %>%
  summarise(
    sum_child = sum(child, na.rm = TRUE),
    sum_adult = sum(adult, na.rm = TRUE),
    sum_elderly = sum(elderly, na.rm = TRUE)
  )

for (c in countries) {
  df_country <- filter(sickness_by_year_df, country == c)
  
  plot_country <- ggplot(df_country, aes(x = year)) +
    geom_line(aes(y = sum_child, color = 'Child')) +
    geom_line(aes(y = sum_adult, color = 'Adult')) +
    geom_line(aes(y = sum_elderly, color = 'Elderly')) +
    labs(y = "No of sick people", title=c)
  
  print(plot_country)
  
  df_country_summary <- df_country %>%
    select(-year) %>%
    summary() %>%
    print
}
```
